Tunable Virtual IMU Frame by Averaging Multiple Non-collocated IMUs

### 2024-09-13 paper/main.tex
Updated paper again according to feedback. Added figures used.

In this article, we show that through a weighted averaging of multiple non-collocated IMUs, we can choose an arbitrary VIMU frame of reference that gives quality-of-life improvement and offering additional benefit of reduced noise and bias drift.

\section{Related Work}

@@

The state propagation using IMU measurements follows the rigid-body kinematics in world fixed frame:
\begin{equation}
\begin{align}
    \dot{\textbf{C}}(t) &= \textbf{C}(t) \bm{\omega}(t)^\wedge, \\
    \dot{\textbf{v}}(t) &= \textbf{C}(t) \textbf{a}(t) - \textbf{g}, \\
    \dot{\textbf{p}}(t) &= \textbf{v}(t).
\end{align}
\end{equation}

@@

\section{Simulation}

@@

\begin{figure}
    \centering
    \includegraphics[width=1.0\linewidth]{figures/simulated_omega.png}
    \caption{Simulated Angular Velocity}
    \label{fig:omega}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=1.0\linewidth]{figures/simulated_accel.png}
    \caption{Simulated Linear Acceleration}
    \label{fig:accel}
\end{figure}
% \begin{figure}
%     \centering
%     \includegraphics[width=0.85\linewidth]{figures/simulated_trajectory.png}
%     \caption{Ground truth Trajectory from Integration}
%     \label{fig:trajectory}
% \end{figure}

@@

We also simulate observations of known landmarks from a monocular camera. This turns the estimation problem into a localization problem such that the error is bounded. For each camera frame, we first sample random observations in the imaging plane as well as depth (between 2 and 10 m). We perform reprojection and transform the resulted 3D points to the inertial frame with the ground-truth pose to generate the corresponding landmark positions. Camera observations arrive at 2 Hz. We selected this low update rate to allow error to propagate for better showing the benefits of averaging.

For the estimator, we implemented the \textit{left-}Invariant Extended Kalman Filter (IEKF) \cite{IEKF} as it is more accurate compared to standard EKF methods. At initialization stage, given all IMU poses, we compute the weighting that places the VIMU frame at the body frame. Using the weights, we also calculate the combined noises and bias drift rates. When new IMU measurements arrive, they are weighted averaged and then fed into the estimator as though we have a single IMU. When camera observations arrive, the 2D observations and 3D landmark positions are used for the correction step.

@@

For performance, we compared results of multiple different IMU configurations summarized in Table \ref{tab:imu_config}. Here symmetric configurations have IMUs placed along the principal axes with an offset of 1 m with random orientation. Assymmetric configurations use IMUs perturbed from the symmetric configuration such that unequal weightings are required to have VIMU frame at origin of body frame. We place the VIMU frame at the body frame for all configurations to give a fair comparison of performance (direct comparison with ground truth). Pose of all IMUs used are summarized in Table \ref{tab:imu_pose}.

@@

\begin{table}[h!]
\centering
\caption{Pose of IMUs}
\label{tab:imu_pose}
\begin{tabular}{ccc}
\toprule
\textbf{IMU ID} & \textbf{orientation} & \textbf{position} \\
\midrule
0      & $1.000 +0.000i +0.000j + 0.000k$ & [$+0.0, +0.0, +0.0$] \\
\midrule
1      & $0.633 +0.065i -0.025j +0.771k$ & [$+1.0, +0.0, +0.0$] \\
2      & $0.395 +0.795i +0.322j +0.328k$ & [$-1.0, +0.0, +0.0$] \\
3      & $0.533 -0.314i +0.671j +0.409k$ & [$+0.0, +1.0, +0.0$] \\
4      & $0.600 +0.323i +0.620j +0.390k$ & [$+0.0, -1.0, +0.0$] \\
5      & $0.231 +0.495i +0.343j -0.764k$ & [$+0.0, +0.0, +1.0$] \\
6      & $0.744 -0.473i +0.355j +0.310k$ & [$+0.0, +0.0, -1.0$] \\
\midrule
11     & $0.045 -0.112i +0.670j +0.733k$ & [$+0.9, +0.0, +0.0$] \\
12     & $0.223 -0.059i +0.929j -0.289k$ & [$-1.2, +0.0, +0.0$] \\
13     & $0.119 +0.962i +0.245j -0.031k$ & [$+0.3, +1.1, +0.0$] \\
14     & $0.605 +0.708i +0.335j +0.141k$ & [$+0.2, -0.7, +0.0$] \\
15     & $0.004 -0.299i +0.894j +0.333k$ & [$-0.1, +0.2, +1.3$] \\
16     & $0.598 -0.094i -0.669j -0.432k$ & [$+0.4, -0.3, -1.1$] \\
\bottomrule
\end{tabular}
\end{table}

@@

\section{Real-World Experiment}

@@

Most public datasets only contain a single IMU. We found that the PennCOSYVIO dataset \cite{penncosyvio} meets our needs of containing data from multiple IMUs. This dataset contains 4 different runs with sensor data from three IMUs and multiple cameras. For our experiment, we chose to only use the VI sensor camera as it was provided with complete information regarding intrinsic, extrinsic, and distortion information. We also only used the left camera of the VI sensor such that it forms the minimal configuration with a single IMU. This is to better expose the impact of IMUs on performance. We select the VI sensor IMU as the main IMU as it is more accurate compared to the IMUs in the two Tango devices. We will be comparing the performance of using only the VI sensor IMU vs. VIMU from averaging all three IMUs.

@@

@@

To demonstrate that the weighting works as expected, we designed four configurations to test: 1) single IMU with weights $[\begin{matrix} 1.0 & 0.0 & 0.0 \end{matrix}]$ (here the first weight is for VI sensor IMU, second weight for top Tango IMU, and the third weight is for bottom Tango IMU), 2) averaged VIMU with weights $[\begin{matrix} 0.33 & 0.33 & 0.33 \end{matrix}]$, 3) VIMU closest to body frame origin with weights $[\begin{matrix} 0.4944 & 0.1546 & 0.3509 \end{matrix}]$ (this places the VIMU at $[\begin{matrix} 0.0186 & -0.0012 & -0.0046 \end{matrix}]$ in the body frame), 4) we also designed a case for which some of the weights go negative $[\begin{matrix} 1.2 & -0.1 & -0.1 \end{matrix}]$ (his places the VIMU outside of the triangle formed by the three IMUs).

@@

\begin{figure}
    \centering
    \includegraphics[width=0.85\linewidth]{figures/posyaw_af.png}
    \caption{Example Trajectory (Run=AF). Black is the ground truth. Blue is with single IMU. Red is with averaged VIMU}
    \label{fig:posyaw_af}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=1.0\linewidth]{figures/error_comp_af_nolabel.png}
    \caption{Example Error Accumulation over Time (Run=AF). Top is rotational error. Bottom is positional error. \\
    Single (Blue) Averaged (Red), Centered (Green), Negative (Yellow) }
    \label{fig:error_af}
\end{figure}

>>>>

\textcolor{red}{In this article, we will show that through a weighted averaging of multiple non-collocated IMUs, we can choose an arbitrary VIMU frame of reference as shown in Figure \ref{fig:vimu_frame}, that gives quality-of-life improvement and offering additional benefit of reduced noise and bias drift. In section \ref{related_work} we will compare different approaches of fusing multiple IMUs and in section \ref{methodology} we will discuss the main idea of how to average multiple non-collocated IMUs. Finally in section \ref{simulation} and \ref{real_experiment} we validated our findings through experiments with simulation and a real dataset.}

\begin{figure}
    \centering
    \includegraphics[width=0.7\linewidth]{figures/fig_frame.png}
    % \caption{VIMU frame $\bm{\mathcal{F}}_v$ and individual IMU sensor frame $\bm{\mathcal{F}}_s$}
    \caption{Multiple IMU as one. Virtual IMU from weighted-averaging multiple IMUs}
    \label{fig:vimu_frame}
\end{figure}

\section{Related Work}\label{related_work}

@@

The state propagation using IMU measurements follows the rigid-body kinematics in world fixed frame:
\begin{equation}
\begin{split}
    \dot{\textbf{C}}(t) &= \textbf{C}(t) \bm{\omega}(t)^\wedge, \\
    \dot{\textbf{v}}(t) &= \textbf{C}(t) \textbf{a}(t) - \textbf{g}, \\
    \dot{\textbf{p}}(t) &= \textbf{v}(t).
\end{split}
\end{equation}

@@

\section{Simulation} \label{simulation}

@@

% \begin{figure}[t!]
%     \centering
%     \includegraphics[width=1.0\linewidth]{figures/simulated_omega.png}
%     \caption{Simulated Angular Velocity}
%     \label{fig:omega}
% \end{figure}

% \begin{figure}[t!]
%     \centering
%     \includegraphics[width=1.0\linewidth]{figures/simulated_accel.png}
%     \caption{Simulated Linear Acceleration}
%     \label{fig:accel}
% \end{figure}

\begin{figure*}[ht]
    \centering
    \begin{subfigure}[b]{0.45\linewidth}
        \centering
        \includegraphics[width=9cm]{figures/simulated_omega.png}
        \caption{Simulated Angular Velocity}
        \label{fig:omega}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\linewidth}
        \centering
        \includegraphics[width=9cm]{figures/simulated_accel.png}
        \caption{Simulated Linear Acceleration}
        \label{fig:accel}
    \end{subfigure}
\caption{Simulated kinematics for ground-truth trajectory generation}
\label{fig:simulation_data}
\end{figure*}

@@

We also simulate observations of known landmarks from a monocular camera at body frame. This turns the estimation problem into a localization problem such that the error is bounded. For each camera frame, we first sample random observations in the imaging plane as well as depth (between 2 and 10 m). We perform reprojection and transform the resulted 3D points to the inertial frame with the ground-truth pose to generate the corresponding landmark positions. Camera observations arrive at 2 Hz. We selected this low update rate to allow error to propagate for better showing the benefits of averaging.

For the estimator, we implemented the \textit{left-}Invariant Extended Kalman Filter (IEKF) \cite{IEKF} as it is more accurate compared to standard EKF methods. At initialization stage, given all IMU poses, we compute the weighting that places the VIMU frame at the body frame, which is also the camera frame. Using the weights, we also calculate the combined noises and bias drift rates. When new IMU measurements arrive, they are weighted averaged and then fed into the estimator as though we have a single IMU. When camera observations arrive, the 2D observations and 3D landmark positions are used for the correction step.

@@

For performance, we compared results of multiple different IMU configurations summarized in Table \ref{tab:imu_config}. Here symmetric configurations have IMUs placed along the principal axes with an offset of 1 m with random orientation. Assymmetric configurations use IMUs perturbed from the symmetric configuration such that unequal weightings are required to have VIMU frame at origin of body frame. We place the VIMU frame at the body frame for all configurations to give a fair comparison of performance (direct comparison with ground truth). Poses of all IMUs used are plotted in Figure \ref{fig:imu_configurations}.

@@

% \begin{table}[h!]
% \centering
% \caption{Pose of IMUs}
% \label{tab:imu_pose}
% \begin{tabular}{ccc}
% \toprule
% \textbf{IMU ID} & \textbf{orientation} & \textbf{position} \\
% \midrule
% 0      & $1.000 +0.000i +0.000j + 0.000k$ & [$+0.0, +0.0, +0.0$] \\
% \midrule
% 1      & $0.633 +0.065i -0.025j +0.771k$ & [$+1.0, +0.0, +0.0$] \\
% 2      & $0.395 +0.795i +0.322j +0.328k$ & [$-1.0, +0.0, +0.0$] \\
% 3      & $0.533 -0.314i +0.671j +0.409k$ & [$+0.0, +1.0, +0.0$] \\
% 4      & $0.600 +0.323i +0.620j +0.390k$ & [$+0.0, -1.0, +0.0$] \\
% 5      & $0.231 +0.495i +0.343j -0.764k$ & [$+0.0, +0.0, +1.0$] \\
% 6      & $0.744 -0.473i +0.355j +0.310k$ & [$+0.0, +0.0, -1.0$] \\
% \midrule
% 11     & $0.045 -0.112i +0.670j +0.733k$ & [$+0.9, +0.0, +0.0$] \\
% 12     & $0.223 -0.059i +0.929j -0.289k$ & [$-1.2, +0.0, +0.0$] \\
% 13     & $0.119 +0.962i +0.245j -0.031k$ & [$+0.3, +1.1, +0.0$] \\
% 14     & $0.605 +0.708i +0.335j +0.141k$ & [$+0.2, -0.7, +0.0$] \\
% 15     & $0.004 -0.299i +0.894j +0.333k$ & [$-0.1, +0.2, +1.3$] \\
% 16     & $0.598 -0.094i -0.669j -0.432k$ & [$+0.4, -0.3, -1.1$] \\
% \bottomrule
% \end{tabular}
% \end{table}

% \begin{figure}
%     \centering
%     \includegraphics[width=0.75\linewidth]{figures/fig_sym_cropped.png}
%     \caption{Caption}
%     \label{fig:enter-label}
% \end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[width=0.75\linewidth]{figures/fig_asym_cropped.png}
%     \caption{Caption}
%     \label{fig:enter-label}
% \end{figure}

\begin{figure*}[ht]
    \centering
    \begin{subfigure}[b]{0.45\linewidth}
        \centering
        \includegraphics[width=6cm]{figures/fig_sym_cropped.png}
        \caption{Symmetrical IMU placement}
        \label{fig:sym_imu}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\linewidth}
        \centering
        \includegraphics[width=6cm]{figures/fig_asym_cropped.png}
        \caption{Asymmetric IMU placement}
        \label{fig:_asym_imu}
    \end{subfigure}
\caption{Poses of IMUs in different configurations }
\label{fig:imu_configurations}
\end{figure*}

@@

\section{Real-World Experiment}\label{real_experiment}

@@

Most public datasets only contain a single IMU. We found that the PennCOSYVIO dataset \cite{penncosyvio} meets our needs of containing data from multiple IMUs. This dataset contains 4 different runs with sensor data from three IMUs and multiple cameras as shown in Figure \ref{fig:sensor_on_rig}. For our experiment, we chose to only use the VI sensor camera as it was provided with complete information regarding intrinsic, extrinsic, and distortion information. We also only used the left camera of the VI sensor such that it forms the minimal configuration with a single IMU. This is to better expose the impact of IMUs on performance. We select the VI sensor IMU as the main IMU as it is more accurate compared to the IMUs in the two Tango devices. We will be comparing the performance of using only the VI sensor IMU vs. VIMU from averaging all three IMUs.

@@

\begin{figure}[h]
\centering
\begin{subfigure}{0.35\textwidth}
\centering
\includegraphics[width = \textwidth]{figures/rig.jpg}
\caption{Sensors}
\label{fig:sensor_on_rig}
\end{subfigure}
\begin{subfigure}{0.12\textwidth}
\centering
\includegraphics[width = \textwidth]{figures/rig_cropped.png}
\caption{Ref. Frames}
\label{fig:rig_coordinate_frame}
\end{subfigure}
\caption{(a) shows the type and location of sensors on the test rig. (b) shows the various frame of reference used in the experiments viewing from the side}
\label{fig:combined}
\end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[width=0.75\linewidth]{figures/rig.jpg}
%     \caption{Caption}
%     \label{fig:enter-label}
% \end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[width=0.75\linewidth]{figures/rig.png}
%     \caption{Caption}
%     \label{fig:enter-label}
% \end{figure}

@@

To demonstrate that the weighting works as expected, we designed four configurations to test: 1) single IMU with weights $[\begin{matrix} 1.0 & 0.0 & 0.0 \end{matrix}]$ (here the first weight is for VI sensor IMU, second weight for top Tango IMU, and the third weight is for bottom Tango IMU), 2) averaged VIMU with weights $[\begin{matrix} 0.33 & 0.33 & 0.33 \end{matrix}]$, 3) VIMU closest to body frame origin with weights $[\begin{matrix} 0.4944 & 0.1546 & 0.3509 \end{matrix}]$ (this places the VIMU at $[\begin{matrix} 0.0186 & -0.0012 & -0.0046 \end{matrix}]$ in the body frame), 4) we also designed a case for which some of the weights go negative $[\begin{matrix} 1.2 & -0.1 & -0.1 \end{matrix}]$ (his places the VIMU outside of the triangle formed by the three IMUs). The various reference frames are as shown in Figure \ref{fig:rig_coordinate_frame}.

@@

% \begin{figure}[ht!]
%     \centering
%     \includegraphics[width=0.85\linewidth]{figures/example_trajectory.png}
%     \caption{Example Trajectory (Run=AF). Black is the ground truth. Blue is with single IMU. Red is with averaged VIMU}
%     \label{fig:posyaw_af}
% \end{figure}

% \begin{figure}[ht!]
%     \centering
%     \includegraphics[width=1.0\linewidth]{figures/example_error.png}
%     \caption{Example Error Accumulation over Time (Run=AF). Top is rotational error. Bottom is positional error. \\
%     Single (Blue) Averaged (Red), Centered (Green), Negative (Yellow) }
%     \label{fig:error_af}
% \end{figure}

\begin{figure*}[ht]
    \centering
    \begin{subfigure}[b]{0.45\linewidth}
        \centering
        % \includegraphics[width=7cm]{figures/posyaw_af.png}
        \includegraphics[width=8cm]{figures/example_trajectory.png}
        \caption{Example Trajectories. All trajectories start at (0, 0).}
        \label{fig:posyaw_af}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\linewidth}
        \centering
        % \includegraphics[width=11cm]{figures/error_comp_af_nolabel.png}
        \includegraphics[width=9cm]{figures/example_error.png}
        \caption{Example errors over Time Rot. (Top), Pos. (Bottom).}
        \label{fig:error_af}
    \end{subfigure}
\caption{Example Result for Run AF. Single (Blue) Averaged (Red), Centered (Green), Negative (Yellow)}
\label{fig:real_example}
\end{figure*}

@@
